---
title: "covid 19"
author: "Janette Quezada"
date: "10/12/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex

$H_0$Dying is independent of sex.
$H_A$Dying depends on sex.

Based on gender we will be using the CAT helper file.
Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```


Among the females 25 out of 1831 died.Among the males 44 out of 1411.According to this the male is most likely to die compare to female.

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

  

Among those who died from covid, according to the data plot chart the males are more likely to die compared to female.



  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

It looks like twice as many more males died than females according to th numerical summary. 

## Inferential results 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value is 0.001, we reject the null hypothesis because it is lower than 0.05.The odds are that males in Korea are 2.3 times more likely to die than females.


## Conclusions

Among those in Korea Males are most likely to die according to the fisher test the p value is 0.0009.In the chi-squared test the p value is  0.0012 which is lower than 0.05 we reject the null hypothesis.Dying is not independent of sex.

